---
title: "Day 4 Notes"
author: "Elise Gonzales"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Nested for loop
```{r}
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for (i in seq_along(odds)) {
  for (j in seq_along(evens)) {
    print(odds[i] * evens[j])
  }
}
```
# inner-most for loop must be evaluated all the way through before outer-most for loop can advance.

## Functions

Anatomy of a function
```{r}
function_name <- function(bird, dog) {
  pets = bird + dog
  return(pets)
}
```
In Python
# def function_name(bird, dog):
  # pets = bird + dog
  # return(pets)
  
```{r}
exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 10)
```

In class warm-up
```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

#you could create empty vector here to store outputs in

for (i in seq_along(quarter_splits)) {
  val = quarter_splits[i] + quarter_splits[i + 1]
  print(val)
}
```

Make a function to double user input.
```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(200) #or
double_this(my_value = 200)

double_this(c(1, 2, 10, 100, 175)) #func is doing iteration for us...
```

Write a func that asks for a user's name and age, and returns an exclamation, like "I'm [name], and I am [age] years old."
```{r}
exclaim_age <- function(your_name, your_age) {
  print(paste("My name is", your_name, "and I am", your_age, "years old."))
}
exclaim_age(your_name = "Elise", your_age = 25)
```

## Function with conditional

```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age* 4.7)
  } else 
    print("Animal must be dog or goat.")
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "pika", human_age = 4)

```

Use tribble to make data frame.
```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)
```
My name is __ & I love __.
```{r}
dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$food))
}
dog_menu(name = "Waffle")
dog_menu(name = "Banjo")

```

## Functions & For Loops
```{r}
# Func iterates over every column in data frame, finds mean. 

#function from scratch (apply col_mean, etc does the same thing)
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

Example...
means_over_columns(df = mtcars)
means_over_columns(df = faithful)

```

