---
title: "Day 4 Notes"
author: "Elise Gonzales"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Nested for loop

```{r}
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

```{r}
odds <- c(1, 3, 5)
evens <- c(2, 4, 6, 8)

for (i in seq_along(odds)) {
  for (j in seq_along(evens)) {
    print(odds[i] * evens[j])
  }
}
```

inner-most for loop must be evaluated all the way through before outer-most for loop can advance.

## Functions

Anatomy of a function

```{r}
function_name <- function(bird, dog) {
  pets = bird + dog
  return(pets)
}
```

In Python \# def function_name(bird, dog): \# pets = bird + dog \# return(pets)

```{r}
exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 10)
```

In class warm-up

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

#you could create empty vector here to store outputs in

for (i in seq_along(quarter_splits)) {
  val = quarter_splits[i] + quarter_splits[i + 1]
  print(val)
}
```

Make a function to double user input.

```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(200) #or
double_this(my_value = 200)

double_this(c(1, 2, 10, 100, 175)) #func is doing iteration for us...
```

Write a func that asks for a user's name and age, and returns an exclamation, like "I'm [name], and I am [age] years old."

```{r}
exclaim_age <- function(your_name, your_age) {
  print(paste("My name is", your_name, "and I am", your_age, "years old."))
}
exclaim_age(your_name = "Elise", your_age = 25)
```

## Function with conditional

```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age* 4.7)
  } else 
    print("Animal must be dog or goat.")
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "pika", human_age = 4)

```

Use tribble to make data frame.

```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)
```

My name is \_\_ & I love \_\_.

```{r}
dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$food))
}
dog_menu(name = "Waffle")
dog_menu(name = "Banjo")

```

## Functions & For Loops

```{r}
# Func iterates over every column in data frame, finds mean. 

#function from scratch (apply col_mean, etc does the same thing)
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

#Example...
means_over_columns(df = mtcars)
means_over_columns(df = faithful)

```

```{r}
apples <- c(1, 2, 3, 4)
price <- c(7, 10, 25)

fruit_out <-matrix(nrow = length(price), ncol = length(apples))

for (i in seq_along(apples)) {
  for (j in seq_along(price)) {
    total_cost <- price[j] * apples[i]
    fruit_out[j, i] <- total_cost #tells you the [row, col] where to place output
  }
}
```

```{r}
first_name <- c("Kiran", "Elise", "Michael")
last_name <- c("Chair", "Table", "Plant", "Desk")

full_name <- matrix(nrow = length(first_name), ncol = length(last_name))

for (i in seq_along(first_name)) {
  for (j in seq_along(last_name)) {
    add_names <- paste(first_name[i], last_name[j])
    full_name[i, j] <- add_names
  }
}
```

## For loop - carrying capacity

Plan & Pseudo-code: 
-- out_matrix = matrix () #to store output 
-- for(i in sequence of growth rates) { 
    for (j in sequence of times) { 
      calc_logistic (for i over all j) 
      store calc_population in out_matrix 
    } 
  }
```{r}
logistic_growth <- function(N0, K, r, time) {
  Nt <- K / (1 + ((K - N0) / N0) * exp(-r * time))
  print(Nt)
}

#iterate over time for multiple growth rates
#Create a seq of growth rate values:
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)

#Need to create a MATRIX to store the outputs in :
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

#Nested for-loop:
for (i in 1:length(r_seq)) { #outer loop is the growth rates
  for (j in 1:length(time_vec)) { #inner loop is the time seq values
  pop <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
  out_matrix[j, i] <- pop #Store the value in the appropriate row, col
  }
  }
```

## Model a population over time and varying growth rate.
```{r}
#Ctrl + Alt + X: creates a function
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

#test function
logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)

#iterate over time
time_vec <- seq(from = 0, to = 35, by = 0.1)

pop_35 <- logistic_growth(N0 = 40, K = 8000, r = 0.1, time = time_vec)


#Data vis
pop_time35_df <- data.frame(time_vec, pop_35)

ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35)) +
         geom_line()
```

same outcome...
```{r}
pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  pop_35_vec[i] <- population
}
```

Extend this to iterate over multiple growth rates.
```{r}
r_seq <- seq(from = 0.20, to = 0.4, by = 0.01)
```

```{r}
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(N0 = 100, K = 4500, r = r_seq[i], time = time_vec[j])
  out_matrix[j, i] <- pop
  }
}

logistic_df <- data.frame(out_matrix, time = time_vec)
colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")

logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate")

ggplot(data = logistic_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate)) +
  theme_classic()
  
```

# Func + warning, error message

Use `warning` to add a warning. Use `stop` to add an error.

$$P=0.3\rho$$
rho: air density (kg / m^3)
A: area covered by turbine (m^2)
v: wind velocity (m/s)
```{r}
calc_wind_power <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    warning("That's fast. Are you sure?")
  }
  
  if (rho > 1.225) {
    warning("That air density is suspicious.")
  }
  
  if (radius < 0) {
    stop("Rotor radius must be a positive value in meters.")
  }
  
  0.3*rho*pi*(radius^2)*(windspeed^3)
}


calc_wind_power(1.225, -6, 15)
```

